### 2010-02-09 doc/manuals/configuration/configuration_manual.tex
Adding more schemas and reducing their font size
Removing HTTP schema from configuration manual.
Added references to another manuals and description of SAML Token SHC

\subsection{Service configuration}\label{sec:Service configuration}
The following services are part of WS-ARC and specific configuration details for
these are given in the respective references below:
\begin{description}
\item[A-REX]
  A-REX
\item[Chelonia]
\item[ISIS]
\item[Charon]
\item[hopi]
  hopi, a lightweight HTTP webserver, is described in details in \%\% REFERENCE.

\end{description}

\subsection{Message Chain}\label{sec:Message Chain}
To configure WS-ARC services, a basic knowledge about how these are
communicating with clients or other services is needed. WS-ARC uses a chain of
message components (MCCs) which serves different levels of functionality. In one
end it accepts messages from an interface and in the other a service processes
them and produce a response. Two types of MCCs exist, and they differ in that
the first type should be able to listen on some interface for messages, while
the other simply
receives messages from another MCC. The MCC then process the message and pass it
on to a single or multiple other MCCs or services. MCCs should be configured
using the XML element \texttt{Component}, where the \texttt{name} attribute
specifies the MCC to use. The following MCCs is part of WS-ARC:
\begin{description}
\item[\texttt{tcp.service}] Listens on the TCP interface, see
  section~\ref{sec:tcp-mcc},
\item[\texttt{tls.service}] Adds encryption layer to communication, see
  section~\ref{sec:tls-mcc},
\item[\texttt{http.service}] Processes HTTP messages, see
  section~\ref{sec:http-mcc},
%~ \item[\texttt{gsi.service}] Processes GSI messages, see
  %~ section~\ref{sec:gsi-mcc},
\item[\texttt{soap.service}] Processes SOAP messages, see
  section~\ref{sec:soap-mcc}.
\end{description}
Additionally a special MCC,
the Plexer, exist which is able to pass messages to different components simply
based on the endpoint the messages was sent to, and it should should be
configured using the \texttt{Plexer} element, see section~\ref{sec:plexer}. In
the end of the chain a service should be specified, using the  \texttt{Service}
element, where the \texttt{name} attribute indicates the given service.

A working configuration should then link a set of MCCs and services, starting
with a MCC that receives messages from some interface. The MCCs and services
should be uniquely identifiable, by setting the \texttt{id} attributes of these
elements. A MCC links to another or several MCCs or services using the
\texttt{next} element, where the \texttt{id} attribute specifies the \texttt{id}
of that MCC or service. The message chain should be contained in the root
element \texttt{ArcConfig}, which is shown in the conceptual configuration
example below:
\begin{lstlisting}[language=xml]
...
<Component name="mcc1.service" id="mcc1">
  <next id="mcc2"/>
</Component>
<Component name="mcc2.service" id="mcc2">
  <next id="service"/>
</Component>
<Service name="service" id="service"/>
...
\end{lstlisting}
For readability the message chain can be grouped into a \texttt{Chain} element
in one or more levels.

@@

\subsubsection{HTTP MCC}\label{sec:http-mcc}
The HTTP protocol is supported by the HTTP MCC named \texttt{http.service}, and
it should either be put on top of the TCP or TLS MCC. A working configuration of
this MCC is simple, since it should only use the \texttt{next} element which
must have a value equal to one of the HTTP request methods. An example of a HTTP
MCC configuration is shown below.

@@

\subsubsection{Plexer}\label{sec:plexer}
The Plexer is a special MCC which filters messages based on their URL endpoints.
One or more regular expressions (POSIX extended) should be defined using one or
more \texttt{next} elements, and these will then be used to filter incoming
messages, thus forwarding messages, where the path of the endpoint URL matched a
given regular expression, to the component or service specified by the
\texttt{id} attribute of the corresponding \texttt{next} element. At the first
match the message will be forwarded to the respective component or service and
no more expressions will be tried. If a message did not match any of the regular
expressions then it is discarded and an error code is returned to the client. An
example of a Plexer configuration is shown below.
%% What is the order testing the regular expressions. First listed first tested or Last listed firsted tested?

@@

@@

\subsubsection{X.509 Token}\label{sec:x509token-shc}
The X.509 Token Security Handler processes information in the SOAP Header
stored according to "Web Services Security X.509 Token Profile 1.0"
\cite{ws-security-x509token}. The plugin name of this Security Handler
is \texttt{x509token.handler} and it resides in \texttt{arcshc} loadable
module.

@@

\subsubsection{SAML Token}

@@

@@

\subsection{TCP Schema}
\lstinputlisting[language=xml]{tcp.xsd}

\subsection{TLS Schema}
\lstinputlisting[language=xml]{tls.xsd}

\bibliography{grid}
\end{document}

>>>>

\subsection{Service configuration}\label{sec:Service configuration}
The following services are part of WS-ARC and specific configuration details for
these are given in the respective references below:
\begin{description}
\item[A-REX] - configuration manual in "The ARC Computational Job Management
Module - A-REX" \cite{a-rex},
\item[Chelonia] storage multi-service - "Chelonia Adminstrator's Manual"
\cite{storage-manual},
\item[ISIS] distributed P2P information service - "ARC Information System"
\cite{isis},
\item[Charon] authorization service - currently described in "Security
Framework of ARC NOX" \cite{arc1-security},
\item[HOPI] lightweight HTTP webserver - "The HOPI Manual" \cite{hopi} 
and also partially described in "Chelonia Administrator's Manual"
\cite{storage-manual}.
\end{description}

\subsection{Message Chain}\label{sec:Message Chain}
To configure WS-ARC services, a basic knowledge about how these are
communicating with clients or other services is needed. WS-ARC uses a chain of
message components (MCCs) which serves different levels of functionality. In one
end it accepts messages from an interface and in the other a service processes
them and produce a response. Two types of MCCs exist, and they differ in that
the first type should be able to listen on some interface for messages, while
the other simply receives messages from another MCC. The MCC then process
the message and pass it on to a single or multiple other MCCs or services.
MCCs should be configured using the XML element \texttt{Component}, where the
\texttt{name} attribute specifies the MCC to use. The following MCCs are part
of WS-ARC:
\begin{description}
\item[\texttt{tcp.service}] Listens on the TCP interface, see
  section~\ref{sec:tcp-mcc},
\item[\texttt{tls.service}] Adds encryption layer to communication, see
  section~\ref{sec:tls-mcc},
\item[\texttt{http.service}] Processes HTTP messages, see
  section~\ref{sec:http-mcc},
\item[\texttt{gsi.service}] Processes GSI messages as implemented in 
Globus toolkit \cite{globus}, see section~\ref{sec:gsi-mcc},
\item[\texttt{soap.service}] Processes SOAP messages, see
  section~\ref{sec:soap-mcc}.
\end{description}
Additionally a special MCC,
the Plexer, exists which is able to pass messages to different components simply
based on the endpoint the messages was sent to, and it should be
configured using the \texttt{Plexer} element, see section~\ref{sec:plexer}. In
the end of the chain a service should be specified, using the \texttt{Service}
element, where the \texttt{name} attribute indicates the given service.

A working configuration should then link a set of MCCs and services, starting
with a MCC that receives messages from some interface. The MCCs and services
should be uniquely identifiable, by setting the \texttt{id} attributes of these
elements. A MCC links to another or several MCCs or services using the
\texttt{next} element, where the \texttt{id} attribute specifies the \texttt{id}
of that MCC or service. Some MCCs support multiple \texttt{next} elements with
textual label inside. Those are used by such MCCs to perform internal routing
of processed message. But most MCCs do not do any routing and hence require
single \texttt{next} without label.

The message chain should be contained in the root element \texttt{ArcConfig},
which is shown in the conceptual configuration example below:
\begin{lstlisting}[language=xml]
...
<Component name="mcc1.service" id="mcc1">
  <next id="mcc2"/>
</Component>
<Component name="mcc2.service" id="mcc2">
  <next id="service"/>
</Component>
<Service name="service" id="service"/>
...
\end{lstlisting}
For readability the message chain can be grouped into a \texttt{Chain} element
in one or more levels.

More technical information about available MCCs and the messages their process
may be found in the Hosting Environment technical manual \cite{hed}.

@@

\subsubsection{HTTP MCC}\label{sec:http-mcc}
The HTTP protocol is supported by the HTTP MCC named \texttt{http.service}, and
it should either be put on top of the TCP or TLS MCC. A working configuration of
this MCC is simple, since it should only use the \texttt{next} element.
Differently from most other MCCs this one may have multiple \texttt{next} with
content being name of HTTP method in uppercase. Depending on method client
requested message may be routed to different components. Most services 
of WS-ARC support GET and PUT for data communication and POST for SOAP. Some
also support HEAD.

An example of a HTTP MCC configuration is shown below.

@@

\subsubsection{Plexer}\label{sec:plexer}
The Plexer is a special MCC which filters messages based on their URL endpoints.
It supports multiple \texttt{next} elements with content being regular expressions
(POSIX extended) used to filter incoming messages.
The message is forwarded to the component or service specified by the \texttt{id}
attribute of the corresponding \texttt{next} element which mathed the endpoint 
URL of the message. At the first match the message will be forwarded to the
respective component or service and no more expressions will be tried. The
\texttt{next} elements are tried in order they are specified in configuration.
If the message did not match any of the regular expressions then it is discarded
and an error is returned to the client. An example of a Plexer configuration is
shown below.

@@

For more technical information about security-related attributes of the
message and the Security Handlers please see \cite{arc1-security}.

@@

\subsubsection{X.509 Token}\label{sec:x509token-shc}
The X.509 Token Security Handler processes information in the SOAP Header
stored according to "Web Services Security: X.509 Token Profile 1.0"
\cite{ws-security-x509token}. The plugin name of this Security Handler
is \texttt{x509token.handler} and it resides in \texttt{arcshc} loadable
module.

@@

\subsubsection{SAML Token}
The SAML Token Security Handler processes information in the SOAP Header
stored according to "Web Services Security: SAML Token Profile 1.1"
\cite{ws-security-samltoken}. The plugin name of this Security Handler
is \texttt{samltoken.handler} and it resides in \texttt{arcshc} loadable
module.

The configuration of this plugin is identical to previously described
X.509 Token Security Handler. It supports configuration elements 
\texttt{Process}, \texttt{CACertificatePath} and \texttt{CACertificatesDir}.

The information collected by this plugin may then be used in ARC Policy 
Decision Point and Service Invoker Policy Decision Point plugins.

@@

 \lstset{
         basicstyle=\footnotesize\ttfamily,
 }

@@

\subsection{TCP MCC Schema}
\lstinputlisting[language=xml]{tcp.xsd}

\subsection{TLS MCC Schema}
\lstinputlisting[language=xml]{tls.xsd}

\subsection{GSI MCC Schema}
\lstinputlisting[language=xml]{gsi.xsd}

\subsection{ARC Authorization Security Handler Schema}
\lstinputlisting[language=xml]{ArcAuthZ.xsd}

\subsection{UsernameToken Security Handler Schema}
\lstinputlisting[language=xml]{UsernameTokenSH.xsd}

\subsection{X.509 Token Security Handler Schema}
\lstinputlisting[language=xml]{X509TokenSH.xsd}

\subsection{SAML Token Security Handler Schema}
\lstinputlisting[language=xml]{SAMLTokenSH.xsd}

\subsection{ARC PDP Schema}
\lstinputlisting[language=xml]{ArcPDP.xsd}

\subsection{XACML PDP Schema}
\lstinputlisting[language=xml]{XACMLPDP.xsd}

\subsection{Delegation PDP Schema}
\lstinputlisting[language=xml]{DelegationPDP.xsd}

\subsection{Service Invoker PDP Schema}
\lstinputlisting[language=xml]{PDPServiceInvoker.xsd}

\subsection{Simple List PDP Schema}
\lstinputlisting[language=xml]{SimpleListPDP.xsd}

\bibliography{grid,nordugrid}
\end{document}
